---
title: "Introduction-to-farsrr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction-to-farsrr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(farsrr)
```


The goal of farsr is to offer tools to report and analyse accidents
reports over years.The farsr package provides functions to read,
summarize and visualize data on accidents recorded by time, date,
geographical information such as: city, state, longitude, latitude… and
more.

You can download and install the development version of farsr using:

  devtools::install_github("hugo7gpad/farsr", build_vignettes = TRUE)


The following functions: make_filename(), fars_read(), fars_read_years(),
fars_summarize_years() and fars_map_state() will be exported once the package is installed.  The 'farsr' requires the installation of
'dplyr', 'tidyr', 'magrittr', 'maps', 'readr', 'tibble' and 'graphics' packages.
An example of data from The US NHTSA Fatality Analysis Reporting System is
included to showcase these functions.

The make_filename() function takes a numeric argument year to create a file name in a specific format for the file that contains all accidents and  related informations reported in a given year.

It can be used in the following way:

```{r make_filename, echo=TRUE}
make_filename(201)
make_filename(2022)
make_filename(2015)
```



The fars_read() function takes a string argument filename and reads its file into R as tibble.

It works like this:

```{r fars_read, echo=TRUE}

```



A list or numeric vector years is the argument to supply to the function fars_read_years(). For each number year in the argument years, a file name is created. fars_read_years() returns a two columns tibble: MONTH, year for each
`year` by adding a new column (year) to the file of that year and selecting the columns MONTH, year. It will return a list of these 2 columns tibbles.

Let's try it!

```{r fars_read_years, echo=TRUE}
yrs2 <- c(2014, 2015)
yrs3 <- c(2013, 2014, 2015)

fars_read_years(2015)
fars_read_years(yrs3)
fars_read_years(yrs2)
```



fars_summarize_years() takes the same argument as fars_read_years(). Its combines all the rows of the list of 2 columns tibbles returned by fars_read_years(). It regroups and summarizes the result to return a new 12 rows tibbles, summary of the numbers of accidents in each month of each element of the `years`.argument supplied.

Examples

```{r fars_summarize_years, echo=TRUE}
yrs2 <- c(2014, 2015)
yrs3 <- c(2013, 2014, 2015)


```



Two numeric arguments(one for state number, the other for year) must be supplied to fars_map_state(). The file associated with the year argument is substted using STATE == state.num . This new data frame is used to draw a map of a region in a state with locations where accidents occured in that region.

```{r fars_map_state, echo=TRUE}

```


In pratical terms, to find how many accidents were reported in every month in the years 2013, 2014, 2015 to the US NHTSA F.A.R.S.' you can run the following R scripts:

```{r probleme, echo=TRUE}
## Accidents reported to the US NHTSA F.A.R.S in 2013, 2014, 2015


# View 2 columns MONTH, year data for each year


# Number of accidents in every month of each year



